\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry}
\geometry{margin=1in}
\title{Information Saturation Theorem}
\author{}
\date{}

\begin{document}
\maketitle

\section*{I. Theorem Statement}

Autonomous systems — both biological and artificial — will experience a decline in performance when information density exceeds a system-specific critical threshold.

Let:
\begin{itemize}
  \item $D(t)$: Information density over time
  \item $P(t)$: System performance
  \item $\theta_i$: Critical information saturation threshold
\end{itemize}

If:
\[
D(t) > \theta_i \quad \Rightarrow \quad \frac{dP}{dt} < 0
\]

Then the system enters a performance degradation phase, reflected by decreased symbolic coherence, accuracy, or interpretability.

\section*{II. Formal Definitions}

\begin{itemize}
  \item \textbf{Information Density} $D(t)$: Volume and entropy of data processed per unit time.
  \item \textbf{Performance Function} $P(t)$: Measurable output quality (accuracy, coherence).
  \item \textbf{Saturation Threshold} $\theta_i$: Maximum density the system can sustain before degradation.
\end{itemize}

\section*{III. Key Mechanisms}

\subsection*{Adaptive Filtering}

Function $F(t)$ dynamically prioritizes relevant data and discards low-entropy inputs to maintain stability near $\theta_i$.

\subsection*{Self-Optimization Protocols}

Feedback systems modify intake and processing behavior to mitigate performance loss near saturation.

\subsection*{Collapse Gradient}

If $D(t)$ continues to exceed $\theta_i$, the degradation of $P(t)$ becomes nonlinear, representing a collapse in coherence.

\section*{IV. Testable Predictions}

\subsection*{AI Systems}

\begin{itemize}
  \item Recursive inputs reduce coherence scores (e.g., BERTScore, BLEU).
  \item Logit entropy increases under saturation.
  \item Hallucination frequency rises in generative models.
\end{itemize}

\subsection*{Human Cognition}

\begin{itemize}
  \item EEG entropy and pupil dilation increase with information overload.
  \item Reaction time and symbolic reasoning accuracy decline.
\end{itemize}

\section*{V. Simulation Design}

\subsection*{Model Setup}

\begin{itemize}
  \item \textbf{Models:} Transformers, VAEs, symbolic logic nets
  \item \textbf{Input:} Increasing entropy and volume data streams
  \item \textbf{Metrics:} Mutual Information, BERTScore, error rate, coherence
\end{itemize}

\subsection*{Cognitive Parallel}

Human subjects under memory and attention constraints demonstrate phase-shift performance collapse.

\section*{VI. Practical Implications}

\begin{itemize}
  \item \textbf{AI Safety:} Avoid collapse through entropy-aware filters
  \item \textbf{UX/Design:} Regulate interface complexity to reduce overload
  \item \textbf{Cognition:} Model and prevent collapse in high-load environments
  \item \textbf{Systems Theory:} Define performance envelopes in symbolic AI
\end{itemize}

\end{document}